Dancing on Water: The Choreography of Sulfur Dioxide Adsorption to Aqueous Surfaces

### 2011-06-30 Methods.tex
Draft 8

All simulations began with an equilibrated cube of 900 \wat~molecules, with sides of length 30\angs. The long axis of each simulation cell (the axis normal to the water surface) was then lengthened to 120\angs, and the systems were further equilibrated for 10 ns. The simulations all employed periodic boundaries to create an ``infinite-slab'' geometry. After equilibrating the neat-\wat~slabs two types of systems were created by introducing \suldiox: a single-\suldiox~system, herein referred to as the ``neat-water'' system, and a saturated \suldiox~system with many gaseous surface and bulk-water \suldiox~molecules.

The low and high concentration simulated systems, ``neat-water'' and saturated water, respectively, were created as follows: the neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations). The single-\suldiox~neat-water system was then evolved for 2 ns to produce an equilibrated starting configuration. Because of the extremely low concentration of \suldiox~in the neat-water system the surface waters behave similarly to a true neat-water air-liquid interface. The neat-water system orientational results shown later in this work reproduce well the results of our previous orientational studies of surface water behavior.\cite{Walker2006b,Hore2008} The saturated system had 22 \suldiox~molecules introduced to the water slab bulk in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water. Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the saturated water slab to simulate an added \suldiox~gas pressure. The saturated system with both bulk and gaseous \suldiox~was then evolved for 2 ns to produce a starting configuration for further saturated simulations.

@@

Equilibrium simulations involved adding \suldiox~to a water slab and equilibrating as outlined above. The neat-water system had a single \suldiox~added to the center of the water box, representing a concentration of 0.06 M. The more concentrated ``saturated'' system consisted of 22 \suldiox~molecules in the bulk corresponding to a concentration 1.35 M. This saturated system was exposed to an additional 50 \suldiox~in the gas phase above the water surface. After equilibration for 2 ns, both the low and high concentration systems were then evolved for a further 10 ns data collection using a time step of 0.5fs, with atomic coordinates recorded every 100 fs.

@@

% need to add in the specifics about the SMD routine, force constants, calculations, etc.
A second set of simulations began with an equilibrated water slab as in the surface equilibrated method above. However, in both the equilibrated neat-water and saturated systems, a single \suldiox~was introduced 20\angs~above the water slab surface, with the sulfur atom tethered to its initial position. The systems were evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations. Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the neat-water and saturated configurations) to guide the \suldiox~down towards a tethered water near the water slab's center of mass by applying a small steering force to the \suldiox-sulfur atom. This steering technique has been previously developed and used to successfully model chemical events.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab. Each of the SMD simulations were performed for a total of 200 ps, using a time step of 1 fs, and taking snapshots of the system every 25 fs. Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-water slab and the saturated slab configurations before steering the \suldiox~towards the water bulk.

@@

	The first portion of the studies involved creating orientational depth-profiles of \suldiox~and \wat~upon exposure and adsorption of the gas. The reference point used for depth calculations is the location of the water surface. A liquid surface, however, is a dynamic boundary that is neither flat nor stationary. To define a water surface location, several previous studies have used the technique of fitting a line shape to the averaged density profile of the water, and extracting interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given time step. Thus, the density profile fitting will capture averaged widths and locations, not instantaneous values. Similarly, the averaged values of location and width will obscure information about any drift or deformations the surface undergoes.

Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation. Figure \ref{fig:density-flaw}a shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work using the original uncorrected coordinates of the system. The water density profile and location (dashed line in figure \ref{fig:density-flaw}a) is produced by averaging the instantaneous density profile at each time step in the simulation over all the time steps. The water profile was then fit to a $\tanh$ function (black line) to extract the position and width parameters of the water surface. The fitted Gibbs dividing surface has a width of 3.77\angs, which is comparable to values reported for similar neat-water systems.\cite{Dang1997,Hore2008} A bulge in the gas-phase ($\leq 52$\angs) side of the water density profile is indicative of the drift of the water slab over the course of the trajectory. Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point.

	To overcome this problem we define the surface location by calculating a reference location at each time step by averaging the positions of the waters contained in the topmost monolayer. This provides us with a consistent and intuitive reference point in the simulations to which we relate our analyses, but does not increase the computational burden.

The location of the water surface is determined by the averaged positions of a set number of the outer-most waters in the simulated aqueous slabs. The number of waters included in the averaging is determined by taking a few issues into account.  First, counting the waters found in the topmost cross-section of the water slab over several time steps indicated between 65-75 waters that established a full monolayer. This was done by a visual inspection of the slab using the VMD MD visualization package.\cite{HUMP96} Alternatively, assuming a spherical model of water with a radius of 2.2\angs~(the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres yielded a similar number of surface water molecules. Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness values. Taking the close-packed model as a maximum number of waters fit into a flat surface, the authors decided to use 70 water molecules for calculating instantaneous water surface locations.

Using this alternative method for finding the outer monolayer location was implemented, and the surface location is plotted as a function of simulation time in figure \ref{fig:density-flaw}b. It is apparent from the surface location plot that the water slab location, and thus the surface location, drifts over the 10 ns, spanning approx. 12 $\AA$. However, the maximum standard deviation of the positions of the waters comprising the surface layer at each time step is only 1.85 $\AA$. Consequently, all depth locations in our analyses are calculated relative to the instantaneous surface location at the corresponding time step of the trajectories.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/density/density-location.png}
		\caption{(A) Density profiles of \wat~(dashed) and \suldiox~(red) from a 10 ns simulation of the neat-water system with a single sulfur dioxide. The \wat~density profile was fit using a hyperbolic tangent function (black) to extract location and width data for the water surface. (B) The instantaneous location of the outer \wat~monolayer was calculated for each simulation time step (blue), and the surface location extracted from the density fitting was also shown for reference (horizontal dashed line). Long simulations of liquid slabs result in drift of the slab location, and the consequent broadening of the water density profile if the surface location is not determined repeatedly throughout the simulation.}
		\label{fig:density-flaw}
	\end{center}
\end{figure}

@@

The molecular orientation of both the \wat~and \suldiox~is a prerequisite for understanding the chemistry occurring during the \suldiox~adsorption process. With the surface location as defined above, the simulated systems were analyzed to characterize the orientation of \wat~and \suldiox~in various environments above, within, and below the aqueous surface region. The two molecules studied are similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane defined by three atoms. A body-fixed frame is defined for both \wat~and \suldiox~as shown in figure \ref{fig:molecular-frame}. In each analysis a space-fixed reference axis is used that corresponds to the long axis of the system's periodic cell normal to the plane of the water surface. The orientational analyses presented herein focus on two angles used to define molecular orientation. The molecular orientation angles $\theta$ and $\phi$ are determined from a set reference axis as shown in figure \ref{fig:water-angles}a.

	The ``tilt'' angle, $\theta$, defines the angle formed between the molecular bisector vector (the molecular z-axis, pointing from the central atom in the direction of the other two atoms) and the positive system reference axis. Thus the value of $\theta$ falls within a range of $[0,\pi]$. An angle of $\cos(\theta)=1$ indicates a molecule with its bisector aligned with the reference axis, while $\cos(\theta)=-1$ results from an anti-aligned configuration. Sample representations of different values of $\cos(\theta)$ are shown in figure \ref{fig:water-angles}b.

	A second angle, $\phi$, defines the molecular ``twist'' for certain configurations of the molecule. $\phi$ is the angle formed between the vector normal to the plane of the molecule (molecular y-axis) and the system reference axis. The values of $\phi$ fall in the interval $[0,\frac{\pi}{2}]$ because of the symmetry of \wat~and \suldiox~molecules with respect to twist about their bisector axes. For values of $\cos(\theta) \approx 0$, $\phi$ provides additional information about whether the molecular orientation is ``flat'' to the surface (e.g. the plane of the molecule is aligned with the plane of the surface), or if it is perpendicular. The values of $\cos(\phi)$ for different molecular orientations are shown in figure \ref{fig:water-angles}c. %However, the symmetry of the \wat~and \suldiox~molecules create equivalence between the two H, or O, atoms (in the \wat~or\suldiox, respectively). $\phi=-\pi$ is equivalent to $\phi=\pi$, and both situations are equivalent to $\phi=0$. Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$. 
	However, values of $\cos(\theta)$ close to 1 or -1 result in an isotropic distribution in $\cos(\phi)$ because of the symmetry of the plane of the surface in directions perpendicular to the reference axis.

@@

%A simple count of hydrogen bonds is telling of the interactions and environments of an adsorbing \suldiox. However, a more sophisticated analysis is needed to fully characterize the geometry of any complexes that form. In this work a graph theoretical method has been used to identify hydrate complexes that form cyclic structures in addition to those that form more simple hydrogen bonding geometries. Previously published works have detailed graph theory for use in molecular systems, and have used graph theoretical methods to describe ice and water clusters, structural energies, as well as in the study of relationships in biological systems.\cite{Huber2007,Shi2005,Radhakrishnan1991,Anick2002} As such, we will not present here a review of the basic graph theoretical details that are well documented in the referenced works. As shown in figure \ref{fig:so2-complex-graph}, the atoms of a \suldiox~and its nearest neighbors are represented as a set of verrtices, with the covalent and hydrogen bonds between them forming the edges of a graph.

>>>>

All simulations began with an equilibrated cube of 900 \wat~molecules, with sides of length 30\angs. The long axis of each simulation cell (the axis normal to the water surface) was then lengthened to 120\angs, and the systems were further equilibrated for 10 ns. The simulations all employed periodic boundaries to create an ``infinite-slab'' geometry. After equilibrating the neat-\wat~slabs two types of systems were created by introducing \suldiox: a single-\suldiox~system, herein referred to as the ``neat-water'' system, and a ``saturated'' \suldiox~system with many gaseous surface and bulk-water \suldiox~molecules.

The low and high concentration simulated systems, ``neat-water'' and ``saturated water'', respectively, were created as follows: the neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations). The single-\suldiox~neat-water system was then evolved for 2 ns to produce an equilibrated starting configuration. Because of the extremely low concentration of \suldiox~in the neat-water system the surface waters behave similarly to a true neat-water air-liquid interface. The neat-water system orientational results shown later in this work reproduce well the results of our previous orientational studies of surface water behavior.\cite{Walker2006b,Hore2008} The saturated system had 22 \suldiox~molecules introduced to the water slab bulk in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water. Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the saturated water slab to simulate an added \suldiox~gas pressure. The saturated system with both bulk and gaseous \suldiox~was then evolved for 2 ns to produce a starting configuration for further saturated simulations.

@@

Equilibrium simulations involved adding \suldiox~to a water slab and equilibrating as outlined above. The neat-water system had a single \suldiox~added to the center of the water box, representing a concentration of 0.06 M. The more concentrated ``saturated'' system consisted of 22 \suldiox~molecules in the bulk corresponding to a concentration 1.35 M. This saturated system was exposed to an additional 50 \suldiox~in the gas phase above the water surface. After equilibration for 2 ns, both the low and high concentration systems were then evolved for a further 10 ns data collection using a time step of 0.5 fs, with atomic coordinates recorded every 100 fs.

@@

% need to add in the specifics about the SMD routine, force constants, calculations, etc.
A second set of simulations began with an equilibrated water slab as in the surface equilibrated method above. However, in both the equilibrated neat-water and saturated systems, a single \suldiox~was introduced 20\angs~above the water slab surface, with the sulfur atom tethered to its initial position. The systems were evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations. Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the neat-water and saturated configurations) to guide the \suldiox~down towards a tethered water near the water slab's center of mass by applying a small steering force to the \suldiox-sulfur atom. This steering technique has been previously developed and used to successfully model chemical events.\cite{Isralewitz2001,Giorgino2011,Bizzarri2011,Strzelecki2009,Patargias2009,Liu2006} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab. Each of the SMD simulations were performed for a total of 200 ps, using a time step of 1 fs, and taking snapshots of the system every 25 fs. Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-water slab and the saturated slab configurations before steering the \suldiox~towards the water bulk.

@@

	The first portion of the studies involved creating orientational depth-profiles in the interfacial region comprised of \suldiox~and \wat~upon exposure and adsorption of the gas. To do this we must define a reference point in thie interfacial region which we refer to here as the water surface location. Recognizing that a liquid surface is a dynamic boundary that is neither flat nor stationary. Several previous studies have used the technique of fitting a line shape to the averaged density profile of the water, and extracting interfacial shape and location parameters to define the water surface location.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given time step. Thus, the density profile fitting will capture averaged widths and locations, not instantaneous values. Similarly, the averaged values of location and width will obscure information about any drift or deformations the surface undergoes. The analysis presented here attempts to retain these subtleties through the use of a ``corrected'' coordinate system.

Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation. Figure \ref{fig:density-flaw}A shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work using the original uncorrected coordinates of the system taken from the raw atomic positions of the molecular dynamics data output. The water density profile and location (thin grey line in Figure \ref{fig:density-flaw}A) is produced by averaging the instantaneous density profile at each time step in the simulation over all the time steps. The water profile was then fit to a $\tanh$ function (black line) to extract the position and width parameters of the water surface. The fitted Gibbs dividing surface has a width of 3.77\angs, which is comparable to values reported for similar neat-water systems.\cite{Dang1997,Hore2008} A bulge in the gas-phase ($\leq 52$\angs) side of the water density profile is indicative of the drift of the water slab over the course of the trajectory. Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point.

	To overcome this problem we define the water surface location by calculating a reference location at each time step by averaging the positions of the waters contained in the topmost monolayer. This provides us with a consistent and intuitive reference point in the simulations to which we relate our analyses, but does not increase the computational burden. The number of waters included in the averaging is determined by taking a few issues into account. First, counting the waters found in the topmost cross-section of the water slab over several time steps indicated between 65-75 waters that established a full monolayer. This was done by a visual inspection of the slab using the VMD MD visualization package.\cite{HUMP96} Alternatively, assuming a spherical model of water with a radius of 2.2\angs~(the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres yielded a similar number of surface water molecules. Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness values. Taking the close-packed model as a maximum number of waters fit into a flat surface, 70 water molecules were used for calculating instantaneous water surface locations.

This method for finding the outer monolayer location was implemented, and the surface location is plotted as a function of simulation time in Figure \ref{fig:density-flaw}B. It is apparent from the surface location plot that the water slab location, and thus the surface location, drifts over the 10 ns, spanning approx. 12 $\AA$. However, the maximum standard deviation of the positions of the waters comprising the surface layer at each time step is only 1.85 $\AA$. Consequently, all depth locations in our analyses are calculated relative to the instantaneous surface location at the corresponding time step of the trajectories.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/density/density-location.png}
		\caption{(A) Density profiles of \wat~(grey) and \suldiox~(red) from a 10 ns simulation of the neat-water system with a single sulfur dioxide. The \wat~density profile was fit using a hyperbolic tangent function (black) to extract location and width data for the water surface. Position values are based on the atomic position data of the water oxygens and \suldiox~sulfurs from the raw molecular dynamics output. (B) The instantaneous location of the outer \wat~monolayer was calculated for each simulation time step (blue), and the surface location extracted from the density fitting was also shown for reference (horizontal dashed line). Long simulations of liquid slabs result in drift of the slab location, and the consequent broadening of the water density profile if the surface location is not determined repeatedly throughout the simulation.}
		\label{fig:density-flaw}
	\end{center}
\end{figure}

@@

Knowing the molecular orientation of both the \wat~and \suldiox~is a prerequisite for understanding the chemistry occurring during the \suldiox~adsorption process. With the surface location as defined above, the simulated systems were analyzed to characterize the orientation of \wat~and \suldiox~in various environments above, within, and below the aqueous surface region. The two molecules studied are similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane defined by three atoms. A body-fixed frame is defined for both \wat~and \suldiox~as shown in Figure \ref{fig:molecular-frame}. In each analysis a space-fixed reference axis is used that corresponds to the long axis of the system's periodic cell normal to the plane of the water surface. The orientational analyses presented herein focus on two angles used to define molecular orientation. The molecular orientation angles $\theta$ and $\phi$ are determined from a set reference axis as shown in Figure \ref{fig:water-angles}A.

	The ``tilt'' angle, $\theta$, defines the angle formed between the molecular bisector vector (the molecular z-axis, pointing from the central atom in the direction of the other two atoms) and the positive system reference axis. Thus the value of $\theta$ falls within a range of $[0,\pi]$. An angle of $\cos(\theta)=1$ indicates a molecule with its bisector aligned with the reference axis, while $\cos(\theta)=-1$ results from an anti-aligned configuration. Sample representations of different values of $\cos(\theta)$ are shown in Figure \ref{fig:water-angles}B.

	A second angle, $\phi$, defines the molecular ``twist'' for certain configurations of the molecule. $\phi$ is the angle formed between the vector normal to the plane of the molecule (molecular y-axis) and the system reference axis. The values of $\phi$ fall in the interval $[0,\frac{\pi}{2}]$ because of the symmetry of \wat~and \suldiox~molecules with respect to twist about their bisector axes. For values of $\cos(\theta) \approx 0$, $\phi$ provides additional information about whether the molecular orientation is ``flat'' to the surface (e.g. the plane of the molecule is aligned with the plane of the surface), or if it is perpendicular. The values of $\cos(\phi)$ for different molecular orientations are shown in Figure \ref{fig:water-angles}C. %However, the symmetry of the \wat~and \suldiox~molecules create equivalence between the two H, or O, atoms (in the \wat~or\suldiox, respectively). $\phi=-\pi$ is equivalent to $\phi=\pi$, and both situations are equivalent to $\phi=0$. Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$. 
	However, values of $\cos(\theta)$ close to 1 or -1 result in an isotropic distribution in $\cos(\phi)$ because of the symmetry of the plane of the surface in directions perpendicular to the reference axis.

@@

%A simple count of hydrogen bonds is telling of the interactions and environments of an adsorbing \suldiox. However, a more sophisticated analysis is needed to fully characterize the geometry of any complexes that form. In this work a graph theoretical method has been used to identify hydrate complexes that form cyclic structures in addition to those that form more simple hydrogen bonding geometries. Previously published works have detailed graph theory for use in molecular systems, and have used graph theoretical methods to describe ice and water clusters, structural energies, as well as in the study of relationships in biological systems.\cite{Huber2007,Shi2005,Radhakrishnan1991,Anick2002} As such, we will not present here a review of the basic graph theoretical details that are well documented in the referenced works. As shown in Figure \ref{fig:so2-complex-graph}, the atoms of a \suldiox~and its nearest neighbors are represented as a set of verrtices, with the covalent and hydrogen bonds between them forming the edges of a graph.

### 2011-06-30 TransitOrientation.tex
Draft 8

	The orientation of \suldiox~molecules throughout the aqueous adsorption process was monitored during the transit SMD simulations. The angles $\theta$ and $\phi$ of the transiting \suldiox~(figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems. The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface, resulting in the 2-dimensional angle and depth-profile histograms shown in figure \ref{fig:so2-transit-angles}. Only the surface into which the transiting \suldiox~was steered is used in the analyses. In all the data sets, the populations of the angle histograms at each distance from the surface were normalized to aid in comparison of regions with differing \suldiox~residence times.

@@

	From its starting position above the water surface, until the \suldiox~moves to within 10\angs~of the surface of both systems, the orientation is isotropic in $\theta$ and $\phi$ (not shown). Isotropic orientation is manifested in the plots as mostly uniform coloration at a given distance from the surface independent of $\cos$ of $\phi$. Near and into the surface the bisector angle $\theta$ becomes more perpendicular to the water interface ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase, consistent with the equilibrium MD simulation results above. At the point when the \suldiox~reaches the surface location (distance $=0$\angs), the bisector is perpendicular in both the neat-water and saturated systems. Figure \ref{fig:so2-surface-angles}a depicts \suldiox~near the water surface with the bisector aligned perpendicular to the interface. In the absence of simulated ionic species that form through \suldiox-\wat~chemistry at the surface, it is clear that the adsorbing \suldiox~in the gas phase takes on a preferred orientation to adsorb on a water surface. The main difference between the neat and saturated water systems is where the point of the transition from isotropic to preferred orientations is found.

  % talk about theta differences
  Comparing in more detail the difference in these two systems upon \suldiox~approach, for the neat-water surface, a transition occurs at approximately 5\angs~from the surface. This coincides with the approximate distance of bonds formed to the outer-most water molecules. Below 5\angs above the surface, the orientation persists and the \suldiox~begins to interact with the topmost surface waters, and orients accordingly. In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approximately 8\angs~above the surface. The layer of adsorbed \suldiox~already present in the saturated system most likely interacts with the transiting \suldiox~molecule. Also, topmost water molecules from the surface move up to a few \angs~inside the adsorbed \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system. It is remarkable that the orientational trend appears so strongly in the $\theta$ plots even with so little data as was collected from the single \suldiox~molecule of each simulation. From onset of orientation above the surface until 10\angs~below, the \suldiox~holds a preferred orientation.

@@

 With a mostly perpendicular bisector angle, it is expected that the values of $\phi$ for \suldiox~would be isotropic relative to the reference axis. This is the case in both systems, with only a few exceptions. In both systems the $\cos(\phi)$ profiles exhibit mostly isotropic distributions above 0\angs, with several regions of lighter coloration interspersed, but without a clearly formed orientational trend. At the neat-water surface and just below, from 0\angs~to -5\angs~in the water phase, the $\theta$ profile broadens down to $\cos(\theta)=0.5$, coinciding with a clear trend in the profile of $\phi$ near $\cos(\phi)=0$ shown in dark red at the bottom of the plot. This corresponds to \suldiox~inclined up to 60 degrees from the surface normal. The result is that \suldiox~in this region shows a preference for one of the S-O bonds to point more towards the aqueous bulk and the other to point away towards the interface. Figure \ref{fig:so2-surface-angles}b depicts the orientation of the \suldiox~slightly below the water surface. This is in contrast to \suldiox~molecules above the surface, oriented more perpendicularly. The behavior is likely due to the interaction between the waters and one of the S-O bonds leading to a higher solvation than above the water surface. As the \suldiox~is solvated by more highly-coordinated bulk water, the S-O bonds experience less equal interaction environments. Baer et al. noted that their force field model for the \suldiox~does not reproduce well the first hydration shell geometries,\cite{Baer2010} so conclusions regarding the specific interactions and hydrate geometries between the \suldiox~and \wat~cannot be made here. It is notable that the same reorientation does not occur as strongly in the saturated system. The presence of the adsorbed \suldiox~layer apparently decreases the reorienting behavior likely because of the disrupting effect the higher \suldiox~concentration has on the water interactions at the surface. %Below this 5\angs~region under the surface, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.

>>>>

	The orientation of \suldiox~molecules throughout the aqueous adsorption process was monitored during the transit SMD simulations. The angles $\theta$ and $\phi$ of the transiting \suldiox~(Figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems. The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface location, resulting in the 2-dimensional angle and depth-profile histograms shown in Figure \ref{fig:so2-transit-angles}. In all the data sets, the populations of the angle histograms at each distance from the surface were normalized to aid in comparison of regions with differing \suldiox~residence times.

@@

	From its starting position above the water surface, until the \suldiox~moves to within 10\angs~of the water surface location of both systems, the orientation is isotropic in $\theta$ and $\phi$ (not shown). Isotropic orientation is manifested in the plots as mostly uniform coloration at a given distance from the surface independent of $\cos$ of $\phi$. Near and into the interfacial region the bisector angle $\theta$ becomes more perpendicular to the water surface ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase, consistent with the equilibrium MD simulation results above. At the point when the \suldiox~reaches the water surface location (distance $=0$\angs), the bisector is perpendicular to the interface in both the neat-water and saturated systems. Figure \ref{fig:so2-surface-angles}a depicts \suldiox~near the water surface with the bisector aligned perpendicular to the interface. In the absence of simulated ionic species that form through \suldiox-\wat~chemistry at the surface, it is clear that the adsorbing \suldiox~in the gas phase takes on a preferred orientation to adsorb on a water surface. The main difference between the neat and saturated water systems is where the point of the transition from isotropic to preferred orientations is found.

  % talk about theta differences
  Comparing in more detail the difference in these two systems upon \suldiox~approach, for the neat-water surface, a transition occurs at approximately 5\angs~from the surface water location. Below 5\angs~above the surface, \suldiox~has a preferred net orientation and is close enough to the water surface that it begins to interact with the topmost surface waters. In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approximately 8\angs~above the surface. The layer of adsorbed \suldiox~already present in the saturated system most likely interacts with the transiting \suldiox~molecule. Also, topmost water molecules from the surface move up to a few \angs~inside the adsorbed \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system. It is remarkable that the orientational trend appears so strongly in the $\theta$ plots even with so little data as was collected from the single \suldiox~molecule of each simulation. From onset of orientation above the surface until 10\angs~below, the \suldiox~holds a preferred orientation.

@@

 With a mostly perpendicular bisector angle, it is expected that the values of $\phi$ for \suldiox~would be isotropic relative to the reference axis. This is the case in both systems, with only a few exceptions. In both systems the $\cos(\phi)$ profiles exhibit mostly isotropic distributions above 0\angs, with several regions of lighter coloration interspersed, but without a clearly formed orientational trend. At the neat-water surface and just below, from 0\angs~to -5\angs~(in the water phase), the $\theta$ profile broadens down to $\cos(\theta)=0.5$, coinciding with a clear trend in the profile of $\phi$ near $\cos(\phi)=0$ shown in dark red at the bottom of the plot. This corresponds to \suldiox~inclined up to 60 degrees from the surface normal. The result is that \suldiox~in this region shows a preference for one of the S-O bonds to point more towards the aqueous bulk and the other to point away towards the interface. Figure \ref{fig:so2-surface-angles}b depicts the orientation of the \suldiox~slightly below the water surface. This is in contrast to \suldiox~molecules above the surface, oriented more perpendicularly. The behavior is likely due to the interaction between the waters and one of the S-O bonds leading to a higher solvation than above the water surface. As the \suldiox~is solvated by more highly-coordinated bulk water, the S-O bonds experience less equal interaction environments. Baer et al. noted that their force field model for the \suldiox~does not reproduce well the first hydration shell geometries,\cite{Baer2010} so conclusions regarding the specific interactions and hydrate geometries between the \suldiox~and \wat~cannot be made here. It is notable that the same reorientation does not occur as strongly in the saturated system. The presence of the adsorbed \suldiox~layer apparently decreases the reorienting behavior likely because of the disrupting effect the higher \suldiox~concentration has on the water interactions at the surface. %Below this 5\angs~region under the surface, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.
